The pragmatics of perspective-taking

### 2018-05-08 writing/journal_manuscript/manuscript.tex
Fill our rest of exp. 1 results, discussion

Even if participants in the unscripted condition make fewer actual errors, they may still be \emph{considering} the hidden object just as often on trials where they go on to make correct responses. As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis. Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult to measure. So we chose as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor. We found that listeners in the scripted condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. In the unscripted condition, however, there was no difference from the baseline, giving rise to a significant interaction in a maximal mixed-effect model with both dyad-level and object-level random effects ($z = -2.7, p = 0.006$; Fig. \ref{fig:exp1results}C). %predicted that even when participants went on to make a correct decision, they would be 
%We define the decision window as the span of time between  their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. 
% Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor. %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. 
%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

@@

Next, we test whether these improvements in listener performance are in fact due to more informative speaker behavior in the unscripted condition. %By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
%First, on filler trials in both conditions, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. For example, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
%Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
%If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
%That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
%We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
We extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances. We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects. \todo[inline]{Update with details once finalized} Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ). In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (see Fig. \ref{fig:exp1results}D), $b = AAA, t = BBB, p < CCC$. The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this scenario.

Does differential informativity of scripted utterances account for the variability across items that we noted in Expt.~1? In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit. We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$. This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances.

After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials? We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

@@

In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it. Furthermore, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell. These results support the claim that listeners neglect the speaker's perspective.

Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.  Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level.

However, when interacting with a natural

@@

In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts (e.g. contexts where objects are occluded such as those used in Director-Matcher tasks to test theory of mind use).

>>>>

Even if participants in the unscripted condition make fewer actual errors, they may still be \emph{considering} the hidden object just as often on trials where they go on to make correct responses. As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we therefore conducted a mouse-tracking analysis. Mouse position is highly correlated with gaze in web browsing \citep{ChenEtAl01_MouseCursor, SpiveyGrosjeanKnoblich05,RoddenEtAl08_EyeMouseCoordinationWebSearch} but precise timing can be difficult and unreliable to measure. 
So we used as our dependent variable the proportion of critical instructions that participants hovered over (or `fixated on') a hidden distractor.
\footnote{We preregistered as our dependent variable the mean \emph{amount of time} spent hovering over hidden distractor, but because this DV was exactly zero for the majority of trials and a large, highly variable number for the rest of trials, we decided post-hoc that the mean was not as meaningful as the binarized version. Still, regression results are unchanged when conducted on log-transformed hover times.} 
We found that listeners in the \emph{scripted} condition hovered over the hidden distractor on over 30\% of critical trials relative to a baseline of 10\% when the hidden object was an unrelated filler, again replicating \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. In the unscripted condition, however, there was no difference from the baseline. We found a significant interaction in a maximal mixed-effect model with dyad-level and object-level random intercepts and a dyad-level random effect for the difference from baseline ($z = -2.8, p = 0.004$; Fig. \ref{fig:exp1results}C). %predicted that even when participants went on to make a correct decision, they would be 
%We define the decision window as the span of time between  their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. 
% Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor. %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. 
%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

@@

Next, we test whether these improvements in listener performance are in fact due to more informative speaker behavior in the unscripted condition. %By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
%First, on filler trials in both conditions, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. For example, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
%Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
%If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
%That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
%We therefore expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.
The simplest measure of speaker informativity is the raw number of   words used in referring expressions. Compared to the scripted referring expressions, speakers in the unscripted condition use significantly more words to refer to critical objects ($b = 0.48, t = 2.5, p < 0.001$ in a mixed-effects regression on difference scores with a fixed intercept and random intercepts for object and dyads.) However, this is a very coarse measure: the short ``pyrex glass'' may be more specific than ``large measuring glass'' despite using fewer words, and more words doesn't necessarily mean more information.

\todo[inline]{Update following 2 paragraphs with details once finalized} 
For a more explicit measure, we extracted the referring expressions generated by speakers in all critical trials and standardized spelling and grammar, yielding 122 unique labels after including scripted utterances. We then recruited XXX judges on Amazon Mechanical Turk to provide ratings for how well these labels fit the target and hidden distractor objects. Responses were given on a slider with endpoints labeled ``not at all'' and ``perfectly.''  Inter-rater reliability was relatively high, with intra-class correlation coefficient of YYY (95\% CI = ZZZ). In a mixed model including random intercepts for raters and items, we found a significant crossover interaction of experiment and referent on mean fitness rating (Fig. \ref{fig:exp1results}D, $b = AAA, t = BBB, p < CCC$). The scripted label fit the hidden distractor just as well or better than the target, but unscripted labels fit the target much better and the hidden distractor much worse. In other words, the scripted labels used in \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} were less informative than speakers normally produce in this scenario.

To what extent does differential informativity of utterances account for the heterogeneity in error? %In Figure \ref{fig:itemLevel}, we compare the item-wise error \% and ratio of target fit to distractor fit. We find that across both experiments, participants have significantly higher error \% on items where the speaker's label fits the distractor better than the target, $b = 0.43, t(14) = 3.97, p = 0.001$, capturing a significant portion of variance, $R^2 = .5, F(1,14) = 0.001$. This suggests that item-wise variability across the two experiments is primarily driven by relative informativity of speaker utterances. 
We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, (\emph{coefficient STAT STAT STAT}) and more likely to make errors when the distractor fit is higher, (\emph{coefficient STAT STAT STAT}). Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = XXX, p < YYY$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

@@

In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it, even on trials where they moved the correct object, matchers were more likely to hover their cursor over the occluded cell. Furthermore, in a novel analysis of utterance informativity, we found that patterns of errors across items are well-predicted by relative fit to the target vs. distractor. Taken together, these results support the claim that listeners neglect the speaker's perspective.

At the same time, however, our \emph{speaker} results suggest a deeper explanation for this phenomenon of perspective-neglect. On filler trials in the scripted condition and all trials in the unscripted condition, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. For example, instead of the scripted ``move the brush one square to the right'' when there is only one brush in their display, speakers instead gave descriptions like ``move the hairbrush with two lines on the handle to the right.'' 
Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment, thus increasing the chance of miscommunication from imprecise descriptions.

@@

In Experiment 2, we more directly test the hypothesis that speakers naturally produce more informative referring expressions in more uncertain contexts where items are occluded. For this experiment, we move to a more systematic stimulus domain. The idiosyncracies we observed among the linguistic items used in Expt. 1 make it difficult to normatively assess pragmatic increases in informativity for a given target. For example, when unscripted participants used an additional modifier (`cassette tape') to describe the target tape in one item, it happened to be highly informative relative to the distractor (a roll of tape). Adding a similar modifier (`hair brush') to describe the target in another item, however, was ineffective because both the target and distractor are hair brushes that are confusable at the subordinate level (one is round and one is flat). Unsurprisingly, then, for
